---
title: "Accident Coordinate Mapping"
author: "Andrew Iraola"
date: "1/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plotting LAT/LONGS for Accident Data from BARD
```{r}
#Loading necessary packages
library(tidyverse)
library(janitor)
library(readxl)
library(maps)
library(ggdark)
```

### Reading in our Dataset
```{r}
ReleasableAccidents2018 <- read_excel("2018BARDCoordinateData.xlsx")

BARD <- ReleasableAccidents2018
```

### Running a Count of NA Values by State
```{r}
#I noticed a lot of "NA" values for the LAT/LONG columns for most states. This query will tell us how many coordinates were successfully recorded for each state out of their total accidents recorded. Ex. Arkansas has 17 LAT/LONG observations to work with, and only 5 NAs. However, for most states, this is not the case. 
BARD %>%
  group_by(State) %>%
  count(Latitude!="NA")
```

### Filtering to Focus Specifically on Florida
```{r}
# I also noticed Florida has a fair amount of data to work with. Out of 607 reported accidents, 54 of them have coordinate data to work with. This is still an abysmal 8 percent of the total, but still... something to work with.
Coordinates <- BARD %>%
  select(State, Latitude, Longitude, AccidentCause1) %>%
  filter(State=="FL")
```

### Reading in our Map Data
```{r}
map_borders_FL <- map_data("state", region = "Florida")
map_borders_MI <- map_data("state", region = "Michigan")
```

# Merging our Two Datasets: Florida Coordinate Data with our Map Data
```{r}
#Let's plot it up!
ggplot() +
  #Plotting Florida borders
  geom_polygon(data=map_borders_FL, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Coordinates, aes(x=Longitude, y=Latitude, color="Individual Accident Recorded")) +
               labs(title="Reported Recreational Boating Accidents in Florida (2018)", 
                    size="Individual Accident",
                    subtitle ="Excludes the 553 for which no data was recorded") + 
  dark_mode(theme_minimal()) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title=element_blank())
```

#### Focusing Now On Only Accidents Where Death Was Recorded
```{r}
#Great Work! If I do say so, myself. However, I'm still bummed about all those missing observations. Turns out, if I filter by accidents that had at least one fatality, the LAT/LONG data is much more robust. Let's do that!
Fatality_Coordinates_FL <- BARD %>%
  select(State, Latitude, Longitude, `Number of Deaths`, AccidentCause1, `PFD Worn?`, `Alcohol Involved?`, `Cause of Death`, `Type of Vessel`, `Operator Education`, `Operator Experience`) %>%
  filter(State=="FL") %>%
  filter(`Number of Deaths`>0)
```

### Graph It Up!
```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Florida borders
  geom_polygon(data=map_borders_FL, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_FL, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=AccidentCause1)) +
               labs(title="Recreational Boating Deaths in Florida (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Florida Deaths Mapped.png")
```
Now for PFDs
```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Florida borders
  geom_polygon(data=map_borders_FL, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_FL, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=`PFD Worn?`)) +
               labs(title="Recreational Boating Deaths in Florida (2018)",
                    subtitle = "PFDs were not worn in 81 percent of all fatalities") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        plot.subtitle = element_text(size=9),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green"))  + 
  # Let's annotate that three death incident
  annotate("text", label = "Three Deaths in\n Capsizing Incident", size=2.7, hjust=0.3, vjust=2.5, color="green", family="Rockwell", x=-87.155, y=30.270) +
  geom_curve(aes(x=-87.155, y=30.270, xend=-87, yend=29.7), arrow = arrow(length =unit(0.2, "cm")),
             size=0.4, color="green", curvature = -0.4) + 
  # Now Lets annotate the PWC Collision
  annotate("text", label = "Two Deaths in\n PWC Collision", size=2.7, hjust=1.5, vjust=0,
           color="green", family= "Rockwell", x=-82.57, y=27.967) + 
  geom_curve(aes(x=-82.57, y=27.967, xend=-83, yend=28), arrow=arrow(length=unit(0.2, "cm")),
             size=0.4, color="green", curvature=0)
#+ ggsave("Florida PFD Deaths Mapped.png")
```


Now, for Alcohol's involvement
```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Florida borders
  geom_polygon(data=map_borders_FL, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_FL, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=`Alcohol Involved?`)) +
               labs(title="Recreational Boating Deaths in Florida (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Florida Deaths Mapped.png")
```
Now, for Cause of Death 
```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Florida borders
  geom_polygon(data=map_borders_FL, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_FL, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=`Cause of Death`)) +
               labs(title="Recreational Boating Deaths in Florida (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) +
        guides(size = guide_legend(order=1),
               color = guide_legend(order = 2))
#+ ggsave("Florida CODs Deaths Mapped.png")
```

Now, for VesselType 
```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Florida borders
  geom_polygon(data=map_borders_FL, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_FL, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=`Type of Vessel`)) +
               labs(title="Recreational Boating Deaths in Florida (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) +
        guides(size = guide_legend(order=1),
               color = guide_legend(order = 2))
#+ ggsave("Florida VesselTypes Deaths Mapped.png")
```

```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Florida borders
  geom_polygon(data=map_borders_FL, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_FL, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=`Operator Experience`)) +
               labs(title="Recreational Boating Deaths in Florida (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) +
        guides(size = guide_legend(order=1),
               color = guide_legend(order = 2))
#+ ggsave("Florida OpExperience Deaths Mapped.png")
```

```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Florida borders
  geom_polygon(data=map_borders_FL, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_FL, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=`Operator Education`)) +
               labs(title="Recreational Boating Deaths in Florida (2018)",
                    caption = "Of the 29 accidents for which operator education status was known,\n     79 percent of them had no formal boating education (n=23)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        plot.caption = element_text(hjust=-0.5),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) +
        guides(size = guide_legend(order=1),
               color = guide_legend(order = 2))
#+ ggsave("Florida OpEducation Deaths Mapped.png")
```

### Now, Let's do the same for Michigan
```{r}
#Now, lets do this for Michigan
Fatality_Coordinates_MI <- BARD %>%
  select(State, Latitude, Longitude, `Number of Deaths`, AccidentEvent1, `PFD Worn?`, `Cause of Death`, `Type of Vessel`, `Operator Experience`, `Operator Education`) %>%
  filter(State=="MI") %>%
  filter(`Number of Deaths`>0)
```

```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Michigan borders
  geom_polygon(data=map_borders_MI, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_MI, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=AccidentEvent1)) +
               labs(title="Recreational Boating Deaths in Michigan (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Michigan Deaths Mapped.png")

# I'm going to put these same color categories in the following graphs that utilize Google Maps. However, I believe the best way to see specific colors is with this black background.
# Might prove more difficult to see green/different similar hues with a green background from the satellite photo
```

```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Michigan borders
  geom_polygon(data=map_borders_MI, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_MI, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=`PFD Worn?`)) +
               labs(title="Recreational Boating Deaths in Michigan (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Michigan PFD Deaths Mapped.png")
```

```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Michigan borders
  geom_polygon(data=map_borders_MI, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_MI, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=`Cause of Death`)) +
               labs(title="Recreational Boating Deaths in Michigan (2018)",
                    subtitle = "Of the 20 cases in which cause of death was known,\ndrownings accounted for 85 percent of all fatalities") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        plot.subtitle =element_text(vjust=3,size=9),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Michigan COD Deaths Mapped.png")
```

```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Michigan borders
  geom_polygon(data=map_borders_MI, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_MI, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=`Type of Vessel`)) +
               labs(title="Recreational Boating Deaths in Michigan (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Michigan VesselType Deaths Mapped.png")
```

```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Michigan borders
  geom_polygon(data=map_borders_MI, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_MI, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=`Operator Experience`)) +
               labs(title="Recreational Boating Deaths in Michigan (2018)",
                    caption = "It is clear from our BARD data that operator experience is not being captured\nin a majority of incidences") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        plot.caption = element_text(hjust=0, size=9),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Michigan OpExperience Deaths Mapped.png")
```

```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Michigan borders
  geom_polygon(data=map_borders_MI, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_MI, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`, color=`Operator Education`)) +
               labs(title="Recreational Boating Deaths in Michigan (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Michigan education Deaths Mapped.png")
```


```{r}
# Now Let's try mapping out the same plot as above, but with county lines. 
library(ggmap)

mibox <- make_bbox(lon = Fatality_Coordinates_MI$Longitude, lat = Fatality_Coordinates_MI$Latitude, f=.1)

mibox
#Now when we gram the map, ggmap will try to fit it into that bounding box. Let's try!
# First let's get the map. By default, it gets it from Google. I want it to be a satellite map
mi_map <- get_map(location=mibox, maptype="terrain", source="google")

# Woah! Looks great!
ggmap(mi_map) + geom_point(data=Fatality_Coordinates_MI, mapping =aes(x=Longitude, y = Latitude, size=`Number of Deaths`, color=`Cause of Death`)) + labs(title="Recreational Boating Deaths\nin Michigan (2018)", caption = "Of the 20 cases in which cause of death was known\ndrownings accounted for 85 percent of all fatalities") +
   theme(text=element_text(family="Rockwell", color="black"),
        plot.title=element_text(hjust=0.0, size=12),
        plot.caption = element_text(hjust=0.6),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) + 
  scale_size_continuous(breaks=c(1,2))
#+ ggsave("Michigan Deaths Mapped2.png")
```

### Now, Let's Try Mapping the Southern Region of the US
```{r}
states <- map_data("state")

Fatality_Coordinates_South <- BARD %>%
  select(State, Latitude, Longitude, `Number of Deaths`, `PFD Worn?`, `Cause of Death`) %>%
filter(State %in% c("SC", "FL", "LA", "GA", "MS", "AL")) %>% 
  filter(`Number of Deaths`>0)

map_borders_south <- subset(states, region %in% c("florida", "south carolina", "georgia", "louisiana", "mississippi", "alabama"))

ggplot(data=map_borders_south) + 
  geom_polygon(aes(x=long, y=lat, group=group), fill = "palegreen", color="black") + coord_fixed(1.3)

ggplot() +
  geom_polygon(data=map_borders_south, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_South, aes(x=Longitude, y=Latitude,  size=`Number of Deaths`)) +
               labs(title="Reported Recreational Boating Deaths\n in Southern Region (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="white"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="white")) 
#+ ggsave("Southern Deaths Mapped.png")
```
I'm pretty happy with this. It gives us a way to display regional Deaths. 


```{r}
#NOW, let's try to map these states via Google Maps
southbox <- make_bbox(lon=Fatality_Coordinates_South$Longitude, lat=Fatality_Coordinates_South$Latitude, f=.1)

south_map <- get_map(location=southbox, maptype = "terrain", source="google")

ggmap(south_map) + geom_point(data=Fatality_Coordinates_South, mapping = aes(x=Longitude, y=Latitude, size=`Number of Deaths`, color=`PFD Worn?`)) +
labs(title="Recreational Boating Deaths in Southern Region (2018)") +
   theme(text=element_text(family="Rockwell", color="black"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) + 
      scale_size_continuous(breaks=c(1,2,3)) +
        guides(size = guide_legend(order=1),
               color = guide_legend(order = 2))
  # +ggsave("Southern Deaths PFD Mapped.png")
```

```{r}
#NOW, let's try to map these states via Google Maps
southbox <- make_bbox(lon=Fatality_Coordinates_South$Longitude, lat=Fatality_Coordinates_South$Latitude, f=.1)

south_map <- get_map(location=southbox, maptype = "terrain", source="google")

ggmap(south_map) + geom_point(data=Fatality_Coordinates_South, mapping = aes(x=Longitude, y=Latitude, size=`Number of Deaths`, color=`Cause of Death`)) +
labs(title="Recreational Boating Deaths in Southern Region (2018)", caption= "Of 121 fatalities in which cause of death was known,\n66% were listed as drownings, 28% were the result of trauma") +
   theme(text=element_text(family="Rockwell", color="black"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
      scale_size_continuous(breaks=c(1,2,3)) +
        guides(size = guide_legend(order=1),
               color = guide_legend(order = 2)) 
  # +ggsave("Southern Deaths COD Mapped.png")
```



### Now, Let's Try Mapping the Whole US
```{r}
states <- map_data("state")

Fatality_Coordinates_ALL <- BARD %>%
  select(State, Latitude, Longitude, `Number of Deaths`, `PFD Worn?`) %>%
  filter(!State %in% c("AK", "AM", "PR", "HI")) %>%
  filter(`Number of Deaths`>0) %>%
  filter(!is.na(`PFD Worn?`))

map_borders_all <- subset(states)

ggplot(data=map_borders_all) + 
  geom_polygon(aes(x=long, y=lat, group=group), fill = "palegreen", color="black") + coord_fixed(1.3)

ggplot() +
  geom_polygon(data=map_borders_all, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_ALL, aes(x=Longitude, y=Latitude, color=`PFD Worn?`)) +
               labs(title="Total Recreational Boating Deaths in the US (2018)",
                    subtitle = "PFDs were not worn in 78 percent of all deaths") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="white"),
        plot.title=element_text(hjust=0.0, size=12),
        plot.subtitle=element_text(hjust=0.0, size=10),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="white")) 
#+ ggsave("Total Number of Deaths Mapped.png")
```

```{r}
#NOW, let's try to map these states via Google Maps
USbox <- make_bbox(lon=Fatality_Coordinates_ALL$Longitude, lat=Fatality_Coordinates_ALL$Latitude, f=.1)

US_map <- get_map(location=USbox, maptype = "terrain", source="google")

ggmap(US_map) + geom_point(data=Fatality_Coordinates_ALL, mapping = aes(x=Longitude, y=Latitude, color=`PFD Worn?`)) +
labs(title="Total Recreational Boating Deaths in US (2018)",
     subtitle = "PFDs were not worn in 78 percent of known cases") +
   theme(text=element_text(family="Rockwell", color="black"),
        plot.title=element_text(hjust=0.0, size=12),
        plot.subtitle=element_text(hjust=0.0, size=10),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
      scale_size_continuous(breaks=c(1,2,3))  
   #+ggsave("US Deaths GoogleMapped.png")
```

```{r}
library(RColorBrewer) 

#NOW, let's try to map these states via Google Maps
USbox <- make_bbox(lon=Fatality_Coordinates_ALL$Longitude, lat=Fatality_Coordinates_ALL$Latitude, f=.1)

US_map <- get_map(location=USbox, maptype = "terrain", source="google")


US_map <- ggmap(US_map, extent="device", legend = "none") + stat_density2d(data=Fatality_Coordinates_ALL, mapping = aes(x=Longitude, y=Latitude, fill=..level.., alpha=..level..), geom = "polygon") 

US_map <- US_map + scale_fill_gradientn(colours = rev(brewer.pal(5, "Spectral")))

#Add the location points, color them red and define round shape
US_map + geom_point(data=Fatality_Coordinates_ALL, aes(x=Longitude, y=Latitude), fill="red", shape=23, alpha=0.8) + guides(size=FALSE, alpha=FALSE) + ggtitle("Heat-Mapping Recreational Boating Fatalities (2018)")
```

```{r}
southbox <- make_bbox(lon=Fatality_Coordinates_South$Longitude, lat=Fatality_Coordinates_South$Latitude, f=.1)

south_map <- get_map(location=southbox, maptype = "terrain", source="google")

south_map <-ggmap(south_map, extent = "device", legend = "none") + stat_density2d(data=Fatality_Coordinates_South, mapping = aes(x=Longitude, y=Latitude, fill=..level.., alpha=..level..), geom = "polygon")

south_map <- south_map + scale_fill_gradientn(colours = rev(brewer.pal(5, "Spectral")))

south_map + geom_point(data=Fatality_Coordinates_South, aes(x=Longitude, y=Latitude), fill="red", shape=23, alpha=0.5) +
  guides(size=FALSE, alpha=FALSE, fill=FALSE) + ggtitle("Heat-Mapping Southern Recreational Boating Fatalities (2018)") +  theme(text=element_text(family="Rockwell", color="black"),
        plot.title=element_text(vjust=0.3, size=12)) + ggsave("HeatMap_South.png")
```

