---
title: "BARD Coordinate Mapping Part 2"
author: "Andrew Iraola"
date: "1/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plotting LAT/LONGS for Accident Data from BARD
```{r}
#Loading necessary packages
library(tidyverse)
library(janitor)
library(readxl)
library(maps)
library(ggdark)
```

### Reading in our Dataset
```{r}
ReleasableAccidents2018 <- read_excel("Coordinate Query.xlsx")

BARD <- ReleasableAccidents2018
```

### Running a Count of NA Values by State
```{r}
#I noticed a lot of "NA" values for the LAT/LONG columns for most states. This query will tell us how many coordinates were successfully recorded for each state out of their total accidents recorded. Ex. Arkansas has 17 LAT/LONG observations to work with, and only 5 NAs. However, for most states, this is not the case. 
BARD %>%
  group_by(State) %>%
  count(LATITUDE!="NA")
```

### Filtering to Focus Specifically on Florida
```{r}
# I also noticed Florida has a fair amount of data to work with. Out of 607 reported accidents, 54 of them have coordinate data to work with. This is still an abysmal 8 percent of the total, but still... something to work with.
Coordinates <- BARD %>%
  select(State, LATITUDE, LONGITUDE, NumberDeaths, CauseofDeath, DeceasedPFDWorn, AlcoholUse) %>%
  filter(State=="FL")
```

### Reading in our Map Data
```{r}
map_borders_FL <- map_data("state", region = "Florida")
map_borders_MI <- map_data("state", region = "Michigan")
```

# Merging our Two Datasets: Florida Coordinate Data with our Map Data
```{r}
#Let's plot it up!
ggplot() +
  #Plotting Florida borders
  geom_polygon(data=map_borders_FL, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Coordinates, aes(x=LONGITUDE, y=LATITUDE, color="Individual Accident Recorded")) +
               labs(title="Reported Recreational Boating Accidents in Florida (2018)", 
                    size="Individual Accident",
                    subtitle ="Excludes the 553 for which no data was recorded") + 
  dark_mode(theme_minimal()) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title=element_blank())
```

#### Focusing Now On Only Accidents Where Death Was Recorded
```{r}
#Great Work! If I do say so, myself. However, I'm still bummed about all those missing observations. Turns out, if I filter by accidents that had at least one fatality, the LAT/LONG data is much more robust. Let's do that!
Fatality_Coordinates_FL <- BARD %>%
  select(State, LATITUDE, LONGITUDE, NumberDeaths, CauseofDeath, DeceasedPFDWorn, AlcoholUse) %>%
  filter(State=="FL") %>%
  filter(NumberDeaths>0)
```

### Graph It Up!
```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Florida borders
  geom_polygon(data=map_borders_FL, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_FL, aes(x=LONGITUDE, y=LATITUDE,  size=NumberDeaths, color=DeceasedPFDWorn)) +
               labs(title="Reported Recreational Boating Fatalities in Florida (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Florida Fatalities by PFD Worn.png")
```

```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Florida borders
  geom_polygon(data=map_borders_FL, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_FL, aes(x=LONGITUDE, y=LATITUDE,  size=NumberDeaths, color=AlcoholUse)) +
               labs(title="Reported Recreational Boating Fatalities in Florida (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Florida Fatalities by Alcohol Involved.png")
```

```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Florida borders
  geom_polygon(data=map_borders_FL, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_FL, aes(x=LONGITUDE, y=LATITUDE,  size=NumberDeaths, color=CauseofDeath)) +
               labs(title="Reported Recreational Boating Fatalities in Florida (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Florida Fatalities by Cause of Death.png")
```
### Now, Let's do the same for Michigan
```{r}
#Now, lets do this for Michigan
Fatality_Coordinates_MI <- BARD %>%
select(State, LATITUDE, LONGITUDE, NumberDeaths, CauseofDeath, DeceasedPFDWorn, AlcoholUse) %>%  filter(State=="MI") %>%
  filter(NumberDeaths>0)
```

```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Michigan borders
  geom_polygon(data=map_borders_MI, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_MI, aes(x=LONGITUDE, y=LATITUDE,  size=NumberDeaths, color=DeceasedPFDWorn)) +
               labs(title="Reported Recreational Boating Fatalities in Michigan (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Michigan Fatalities Mapped.png")

# I'm going to put these same color categories in the following graphs that utilize Google Maps. However, I believe the best way to see specific colors is with this black background.
# Might prove more difficult to see green/different similar hues with a green background from the satellite photo
```


```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Michigan borders
  geom_polygon(data=map_borders_MI, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_MI, aes(x=LONGITUDE, y=LATITUDE,  size=NumberDeaths, color=AlcoholUse)) +
               labs(title="Reported Recreational Boating Fatalities in Michigan (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Michigan Fatalities Mapped.png")
```

```{r}
#Let's plot it up!
  ggplot() +
  #Plotting Michigan borders
  geom_polygon(data=map_borders_MI, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_MI, aes(x=LONGITUDE, y=LATITUDE,  size=NumberDeaths, color=CauseofDeath)) +
               labs(title="Reported Recreational Boating Fatalities in Michigan (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2)) + 
  theme(text=element_text(family="Rockwell", color="green"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="green")) 
#+ ggsave("Michigan Fatalities Mapped.png")
```

```{r}
# Now Let's try mapping out the same plot as above, but with county lines. 
library(ggmap)

mibox <- make_bbox(lon = Fatality_Coordinates_MI$LONGITUDE, lat = Fatality_Coordinates_MI$LATITUDE, f=.1)

mibox
#Now when we gram the map, ggmap will try to fit it into that bounding box. Let's try!
# First let's get the map. By default, it gets it from Google. I want it to be a satellite map
mi_map <- get_map(location=mibox, maptype="terrain", source="google")


# Woah! Looks great!
ggmap(mi_map) + geom_point(data=Fatality_Coordinates_MI, mapping =aes(x=LONGITUDE, y = LATITUDE, color=CauseofDeath)) + labs(title="Reported Recreational Boating\n Fatalities in Michigan (2018)") +
   theme(text=element_text(family="Rockwell", color="black"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
#+ ggsave("Michigan Fatalities by Cause of Death.png")
```

### Now, Let's Try Mapping the Southern Region of the US
```{r}
states <- map_data("state")

Fatality_Coordinates_South <- BARD %>%
  select(State, LATITUDE, LONGITUDE, NumberDeaths, DeceasedPFDWorn, AlcoholUse, CauseofDeath) %>%
filter(State %in% c("SC", "FL", "LA", "GA", "MS", "AL")) %>% 
filter(NumberDeaths>0)

map_borders_south <- subset(states, region %in% c("florida", "south carolina", "georgia", "louisiana", "mississippi", "alabama"))

ggplot(data=map_borders_south) + 
  geom_polygon(aes(x=long, y=lat, group=group), fill = "palegreen", color="black") + coord_fixed(1.3)

ggplot() +
  geom_polygon(data=map_borders_south, aes(x=long, y=lat, group = group), color="black", fill="#303030", size=1.15) + 
  #Plotting Accident Coordinates Observed
  geom_point(data=Fatality_Coordinates_South, aes(x=LONGITUDE, y=LATITUDE,  size=NumberDeaths, color=DeceasedPFDWorn)) +
               labs(title="Reported Recreational Boating Fatalities\n in Southern Region (2018)") + 
  dark_mode(theme_minimal()) + 
  scale_size_continuous(breaks=c(1,2,3)) + 
  theme(text=element_text(family="Rockwell", color="white"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.title = element_text(color="white")) 
#+ ggsave("Southern Fatalities by PFD worn.png")
```


```{r}
#NOW, let's try to map these states via Google Maps
southbox <- make_bbox(lon=Fatality_Coordinates_South$LONGITUDE, lat=Fatality_Coordinates_South$LATITUDE, f=.1)

south_map <- get_map(location=southbox, maptype = "terrain", source="google")

ggmap(south_map) + geom_point(data=Fatality_Coordinates_South, mapping = aes(x=LONGITUDE, y=LATITUDE, size=NumberDeaths, color=DeceasedPFDWorn)) +
labs(title="Reported Recreational Boating Fatalities\n in Southern Region (2018)", 
     caption = "Out of 117 fatalities for which PFD wear-status was known,\n only 20 percent of victims were wearing them") +
   theme(text=element_text(family="Rockwell", color="black"),
        plot.title=element_text(hjust=0.0, size=12),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
      scale_size_continuous(breaks=c(1,2,3))  
   #+ggsave("Southern Fatalities by PFD Worn Google.png")
```



