---
title: "Table 1_2019 Redux"
author: "Andrew Iraola"
date: "9/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Annual RBS Statistics: Executive Summary Breakdown (2019)

```{r}
library(readxl)
library(ggplot2)
library(dplyr)

Overview_and_Vessel <- read_excel("table 1 redux.xlsx")
```
Notes:
1. This analysis will exclude categories "Unknown" and "Other" as they are not definitively "Motorized", "Non-Motorized", or "Sail" (which will hereby encompass "Sail (only)", "Sail (unknown propulsion) and "Auxiliary sail" )
2. Houseboat is herein defined as motorized. Though this could be changed if desired. 
```{r}
Motorized_Table <- Overview_and_Vessel %>%
  filter(VesselType=="Airboat"|VesselType=="Cabin motorboat"|VesselType=="Houseboat"|VesselType=="open motorboat"|VesselType=="Open motorboat"|VesselType=="Open Motorboat"|VesselType=="Personal watercraft"|VesselType=="pontoon"|VesselType=="Pontoon") 
```

3. Decided that inflatables are non-motorized. However, we can change this anytime. 
```{r}
NonMotorized_Table <- Overview_and_Vessel %>%
  filter(VesselType=="Canoe"|VesselType=="Inflatable"|VesselType=="Kayak"|VesselType=="Rowboat"|VesselType=="Standup paddleboard"|VesselType=="Standup Paddleboard")
```

```{r}
Sail_Table <- Overview_and_Vessel %>%
  filter(VesselType=="Auxiliary sail"|VesselType=="Sail (unknown propulsion)"|VesselType=="Sail (only)")
```

## 1. a. Let's begin by looking at motorized vessel accident causes
```{r}
df_MotorizedIncident1 <- Motorized_Table %>%
  group_by(VesselCause1) %>%
  summarise(counts = n()) %>%
  filter(VesselCause1!="None; not at fault") %>%
  filter(counts >=150)

ggplot(df_MotorizedIncident1, aes(x = VesselCause1, y = counts)) +
  geom_bar(aes(reorder(VesselCause1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Accident Causes for Motorized Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0,600)) 

# Now, lets look at deaths 
df_MotorizedIncident2 <- Motorized_Table %>%
  group_by(VesselCause1) %>%
   filter(NumberDeaths>0) %>%
summarise(counts = sum(NumberDeaths)) %>%
  filter(counts>=12)
 
ggplot(df_MotorizedIncident2, aes(x = VesselCause1, y = counts)) +
  geom_bar(aes(reorder(VesselCause1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Fatality Causes for Motorized Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0,100)) 

# Now, lets look at injuries 
df_MotorizedIncident3 <- Motorized_Table %>%
  group_by(VesselCause1) %>%
   filter(NumberInjured>0) %>%
   summarise(counts = sum(NumberInjured)) %>%
   filter(counts>=55)
 
ggplot(df_MotorizedIncident3, aes(x = VesselCause1, y = counts)) +
  geom_bar(aes(reorder(VesselCause1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Injury Causes for Motorized Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0,325)) 
```

## 1. b. Now, let's look at Motorized Events
```{r}
df_MotorizedIncident4 <- Motorized_Table %>%
  group_by(VesselEvent1) %>%
  summarise(counts = n()) %>%
  filter(counts>=129)

ggplot(df_MotorizedIncident4, aes(x = VesselEvent1, y = counts)) +
  geom_bar(aes(reorder(VesselEvent1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top 10 Accident Events for Motorized Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0,1500))

# Now, lets look at deaths 
df_MotorizedIncident5 <- Motorized_Table %>%
  group_by(VesselEvent1) %>%
   filter(NumberDeaths>0) %>%
   summarise(counts = sum(NumberDeaths)) 
 
ggplot(df_MotorizedIncident5, aes(x = VesselEvent1, y = counts)) +
  geom_bar(aes(reorder(VesselEvent1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top 10 Fatality Events for Motorized Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  +
  scale_y_continuous(limits = c(0,150))

# Now, lets look at injuries 
df_MotorizedIncident6 <- Motorized_Table %>%
  group_by(VesselEvent1) %>%
   filter(NumberInjured>0) %>%
summarise(counts = sum(NumberInjured))
 
ggplot(df_MotorizedIncident6, aes(x = VesselEvent1, y = counts)) +
  geom_bar(aes(reorder(VesselEvent1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top 10 Injury Events for Motorized Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  +
  scale_y_continuous(limits = c(0,500))
```

## 1. c. Lastly, lets get a handle on the numbers
```{r}
# Lets count up accidents, injuries, deaths, and drownings
Motorized_Table %>%
  count(NumberDeaths) 

Motorized_Table %>%
  count(NumberDrownings)

Motorized_Table %>%
  count(NumberInjured)

Motorized_Table %>%
  count(NumberNonDrownings)
```

# 2. a. Let's begin by looking at non-motorized accident causes
```{r}
df_NonMotorizedIncident1 <- NonMotorized_Table %>%
  group_by(VesselCause1) %>%
  summarise(counts = n()) %>%
  filter(VesselCause1!="None; not at fault") %>%
  filter(counts>=0)

ggplot(df_NonMotorizedIncident1, aes(x = VesselCause1, y = counts)) +
  geom_bar(aes(reorder(VesselCause1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Accident Causes for Non-Motorized Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0,100)) 

# Now, lets look at deaths 
df_NonMotorizedIncident2 <- NonMotorized_Table %>%
  group_by(VesselCause1) %>%
   filter(NumberDeaths>0) %>%
 summarise(counts = sum(NumberDeaths)) %>%
  filter(counts>=0)
 
ggplot(df_NonMotorizedIncident2, aes(x = VesselCause1, y = counts)) +
  geom_bar(aes(reorder(VesselCause1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Fatality Causes for Non-Motorized Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0,50)) 

# Now, lets look at injuries 
df_NonMotorizedIncident3 <- NonMotorized_Table %>%
  group_by(VesselCause1) %>%
   filter(NumberInjured>0) %>%
   summarise(counts = sum(NumberInjured)) %>%
   filter(counts>=0)
 
ggplot(df_NonMotorizedIncident3, aes(x = VesselCause1, y = counts)) +
  geom_bar(aes(reorder(VesselCause1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Injury Causes for Non-Motorized Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0,50)) 
```

# 2. b. Now, lets look at the main events during accidents
```{r}
df_NonMotorizedIncident4 <- NonMotorized_Table %>%
  group_by(VesselEvent1) %>%
  summarise(counts = n()) %>%
  filter(counts>=0)

ggplot(df_NonMotorizedIncident4, aes(x = VesselEvent1, y = counts)) +
  geom_bar(aes(reorder(VesselEvent1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Accident Events for Non-Motorized Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0,150))

# Now, lets look at deaths 
df_NonMotorizedIncident5 <- NonMotorized_Table %>%
  group_by(VesselEvent1) %>%
   filter(NumberDeaths>0) %>%
summarise(counts = sum(NumberDeaths)) 
 
ggplot(df_NonMotorizedIncident5, aes(x = VesselEvent1, y = counts)) +
  geom_bar(aes(reorder(VesselEvent1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Fatality Events for Non-Motorized Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  +
  scale_y_continuous(limits = c(0,100))

# Now, lets look at injuries 
df_NonMotorizedIncident6 <- NonMotorized_Table %>%
  group_by(VesselEvent1) %>%
   filter(NumberInjured>0) %>%
   summarise(counts = sum(NumberInjured)) 
 
ggplot(df_NonMotorizedIncident6, aes(x = VesselEvent1, y = counts)) +
  geom_bar(aes(reorder(VesselEvent1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Injury Events for Non-Motorized Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  +
  scale_y_continuous(limits = c(0,100))
```

# 2. c. Lastly, lets get a handle on the numbers
```{r}
# Lets count up accidents, injuries, deaths, and drownings
NonMotorized_Table %>%
  count(NumberDeaths) 

NonMotorized_Table %>%
  count(NumberDrownings)

NonMotorized_Table %>%
  count(NumberInjured)

NonMotorized_Table %>%
  count(NumberNonDrownings)
```

# 3. a. Let's begin by looking at Sail vessel accident causes
```{r}
df_SailIncident1 <- Sail_Table %>%
  group_by(VesselCause1) %>%
  summarise(counts = n()) %>%
  filter(VesselCause1!="None; not at fault") %>%
  filter(counts >=0)

ggplot(df_SailIncident1, aes(x = VesselCause1, y = counts)) +
  geom_bar(aes(reorder(VesselCause1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Accident Causes for Sail Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0,50)) 

# Now, lets look at deaths 
df_SailIncident2 <- Sail_Table %>%
  group_by(VesselCause1) %>%
   filter(NumberDeaths>0) %>%
summarise(counts = sum(NumberDeaths)) %>%
filter(counts>=0)
 
ggplot(df_SailIncident2, aes(x = VesselCause1, y = counts)) +
  geom_bar(aes(reorder(VesselCause1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Fatality Causes for Sail Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0,10)) 

# Now, lets look at injuries 
df_SailIncident3 <- Sail_Table %>%
  group_by(VesselCause1) %>%
   filter(NumberInjured>0) %>%
   summarise(counts = sum(NumberInjured)) %>%
   filter(counts>=0)
 
ggplot(df_SailIncident3, aes(x = VesselCause1, y = counts)) +
  geom_bar(aes(reorder(VesselCause1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Injury Causes for Sail Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0,15)) 
```

## 3. b. Now, let's look at Sail Events
```{r}
df_SailIncident4 <- Sail_Table %>%
  group_by(VesselEvent1) %>%
  summarise(counts = n()) %>%
  filter(counts>=0)

ggplot(df_SailIncident4, aes(x = VesselEvent1, y = counts)) +
  geom_bar(aes(reorder(VesselEvent1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Accident Events for Sail Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  +
  scale_y_continuous(limits = c(0,70))

# Now, lets look at deaths 
df_SailIncident5 <- Sail_Table %>%
  group_by(VesselEvent1) %>%
   filter(NumberDeaths>0) %>%
   summarise(counts = sum(NumberDeaths))
 
ggplot(df_SailIncident5, aes(x = VesselEvent1, y = counts)) +
  geom_bar(aes(reorder(VesselEvent1, -counts), counts),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Fatality Events for Sail Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  +
  scale_y_continuous(limits = c(0,10))

# Now, lets look at injuries 
df_SailIncident6 <- Sail_Table %>%
  group_by(VesselEvent1) %>%
   filter(NumberInjured>0) %>%
   summarise(Total = sum(NumberInjured)) 
 
ggplot(df_SailIncident6, aes(x = VesselEvent1, y = Total)) +
  geom_bar(aes(reorder(VesselEvent1, -Total), Total),fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = Total), vjust = -0.3) +
  labs(y = "Total Incidents", fill="test", x = "", title = "Top Injury Events for Sail Vessels") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  +
  scale_y_continuous(limits = c(0,20))
```

# 3. c. Lastly, lets get a handle on the numbers
```{r}
# Lets count up accidents, injuries, deaths, and drownings
Sail_Table %>%
  count(NumberDeaths) 

Sail_Table %>%
  count(NumberDrownings)

Sail_Table %>%
  count(NumberInjured)

Sail_Table %>%
  count(NumberNonDrownings)
```

