---
title: "Annual Fire/Explosion Analysis (2019)"
author: "Andrew Iraola"
date: "2/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Annual Fire/Explosion Analysis

Let's begin by loading in the necessary R packages. Then, we'll apply filters to divide our two groups: 
Boats under 20', and boats from 20'-26'
```{r}
library(tidyverse)
library(readxl)
library(dplyr)

Vessel_table <- read_excel("2019_Vessel.xlsx")

Boats_lessthan20 <- Vessel_table %>%
  filter(Length<20, EngineType=="Outboard")

Boats_20to26 <- Vessel_table %>%
  filter(Length>=20,Length<=26, EngineType=="Outboard")
```

## Boats Less Than 20 feet in Length
### Fire/Explosion (fuel) Related
First, let's wrangle any cases that have `Fire/explosion (fuel)` listed as the first, second, or third `VesselEvent`. We'll name this group, `Fuel_related_under20`
```{r}
Fuel_related1 <- Boats_lessthan20 %>%
  filter(VesselEvent1=="Fire/explosion (fuel)") 

Fuel_related2 <- Boats_lessthan20 %>%
  filter(VesselEvent2=="Fire/explosion (fuel)")

Fuel_related3 <- Boats_lessthan20 %>%
  filter(VesselEvent3=="Fire/explosion (fuel)")

Fuel_related_under20 <- rbind(Fuel_related1, Fuel_related2, Fuel_related3)
```

Taking a look at the "Environment" window tells us how many total fire/explosion cases there were related to fuel during this period. 

Now, lets generate some basic descriptive statistics
```{r}
Fuel_related_under20 %>%
  group_by(VesselCause1, VesselCause2, VesselCause3) %>%
  tally
# While this function isn't perfect, it does allow us to easily count up how often specific causes come up across several pages of incidents.
```


```{r}
# This is another way to accomplish the above code chunk. Count how many instances each VesselCause comes up
df_VesselCause1_fuelunder20<- Fuel_related_under20 %>%
  group_by(VesselCause1) %>%
  summarise(counts = n()) 

df_VesselCause2_fuelunder20<- Fuel_related_under20 %>%
  group_by(VesselCause2) %>%
  summarise(counts = n()) 

df_VesselCause3_fuelunder20<- Fuel_related_under20 %>%
  group_by(VesselCause3) %>%
  summarise(counts = n()) 

# If you summarize the values for each of the three dataframes above, you can get to how many total causes are mentioned for either cause1, cause2, or cause3

```

```{r}
df_VesselCause1_fuelunder20 <- Fuel_related_under20 %>%
  select(VesselCause1, VesselCause2, VesselCause3)

df_VesselCause2_fuelunder20 <- Fuel_related_under20 %>%
  select(VesselCause1,VesselCause2, VesselCause3) %>%
  mutate("VesselCause1"=VesselCause2)

df_VesselCause3_fuelunder20 <- Fuel_related_under20 %>%
  select(VesselCause1,VesselCause3, VesselCause2) %>%
  mutate("VesselCause1"=VesselCause3)

Fuel_related_under20_CauseTally <- rbind(df_VesselCause1_fuelunder20, df_VesselCause2_fuelunder20, df_VesselCause3_fuelunder20)

Fuel_related_under20_CauseTally1 <- Fuel_related_under20_CauseTally %>%
  group_by(VesselCause1) %>%
  filter(VesselCause1!="na") %>%
  summarise(counts = n())

# Now, lets graph it up! Finally..
ggplot(Fuel_related_under20_CauseTally1, aes(x = VesselCause1, y= counts )) +
  geom_bar(aes(reorder(VesselCause1, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Incident Cause", title = "Number of Fire/Explosion (Fuel) Incidents by Causes Listed", subtitle = "For vessels under 20 feet") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 15)) 
```

Now, machinery and equipment failures
```{r}
df_VesselMachineryFailure_fuelunder20 <- Fuel_related_under20 %>%
  group_by(VesselMachineryFailure) %>%
  filter(VesselMachineryFailure!="na") %>%
  summarise(counts = n()) 

ggplot(df_VesselMachineryFailure_fuelunder20, aes(x = VesselMachineryFailure, y = counts)) +
  geom_bar(aes(reorder(VesselMachineryFailure, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Machinery Failure (Type)", title = "Number of Fire/Explosion (Fuel) Incidents by Machinery Failure Reported", subtitle = "For vessels under 20 feet") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 10))

################################################################################################

df_VesselEquipmentFailure_fuelunder20 <- Fuel_related_under20 %>%
  group_by(VesselEquipmentFailure) %>%
  filter(VesselEquipmentFailure!="na") %>%
  summarise(counts = n()) 
#As we can see, there were no incidents, where a fire extinguisher failed..
```

What lengths under 20 feet, had fire/explosion related accidents?
```{r}
df_VesselLength_fuelunder20<- Fuel_related_under20 %>%
  group_by(Length) %>%
  summarise(counts = n()) 

ggplot(df_VesselLength_fuelunder20, aes(x = Length, y = counts)) +
  geom_bar(aes(reorder(Length, counts), counts), fill = "dodgerblue", stat = "identity") +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Length", title = "Number of Fire/Explosion (Fuel) Incidents by Vessel Length", subtitle = "For vessels under 20 feet") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 20))
```

###Fire/Explosion (non-fuel) Related
```{r}
NonFuel_related1 <- Boats_lessthan20 %>%
  filter(VesselEvent1=="Fire/explosion (non-fuel)") 

NonFuel_related2 <- Boats_lessthan20 %>%
  filter(VesselEvent2=="Fire/explosion (non-fuel)")

NonFuel_related3 <- Boats_lessthan20 %>%
  filter(VesselEvent3=="Fire/explosion (non-fuel)")

NonFuel_related_under20 <- rbind(NonFuel_related1, NonFuel_related2, NonFuel_related3)


df_VesselCause1_Nonfuelunder20 <- NonFuel_related_under20 %>%
  select(VesselCause1, VesselCause2, VesselCause3)

df_VesselCause2_Nonfuelunder20 <- NonFuel_related_under20 %>%
  select(VesselCause1,VesselCause2, VesselCause3) %>%
  mutate("VesselCause1"=VesselCause2)

df_VesselCause3_Nonfuelunder20 <- NonFuel_related_under20 %>%
  select(VesselCause1,VesselCause3, VesselCause2) %>%
  mutate("VesselCause1"=VesselCause3)

NonFuel_related_under20_CauseTally <- rbind(df_VesselCause1_Nonfuelunder20, df_VesselCause2_Nonfuelunder20, df_VesselCause3_Nonfuelunder20)

NonFuel_related_under20_CauseTally1 <- NonFuel_related_under20_CauseTally %>%
  group_by(VesselCause1) %>%
  filter(VesselCause1!="na") %>%
  summarise(counts = n())

# Now, lets graph it up! Finally..
ggplot(NonFuel_related_under20_CauseTally1, aes(x = VesselCause1, y= counts )) +
  geom_bar(aes(reorder(VesselCause1, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Incident Cause", title = "Number of Fire/Explosion (Non-Fuel) Incidents by Causes Listed", subtitle = "For vessels under 20 feet") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 10)) 
```

Now, machinery and equipment failures
```{r}
df_VesselMachineryFailure_Nonfuelunder20 <- NonFuel_related_under20 %>%
  group_by(VesselMachineryFailure) %>%
  filter(VesselMachineryFailure!="na") %>%
  summarise(counts = n()) 

ggplot(df_VesselMachineryFailure_Nonfuelunder20, aes(x = VesselMachineryFailure, y = counts)) +
  geom_bar(aes(reorder(VesselMachineryFailure, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Machinery Failure (Type)", title = "Number of Fire/Explosion (Non-Fuel) Incidents by Machinery Failure Reported", subtitle = "For vessels under 20 feet") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 8))

################################################################################################

df_VesselEquipmentFailure_Nonfuelunder20 <- NonFuel_related_under20 %>%
  group_by(VesselEquipmentFailure) %>%
  filter(VesselEquipmentFailure!="na") %>%
  summarise(counts = n()) 
#As we can see, there were no incidents, where a fire extinguisher failed..
```

What lengths under 20 feet, had fire/explosion (non-fuel) related accidents?
```{r}
df_VesselLength_Nonfuelunder20<- NonFuel_related_under20 %>%
  group_by(Length) %>%
  summarise(counts = n()) 

ggplot(df_VesselLength_Nonfuelunder20, aes(x = Length, y = counts)) +
  geom_bar(aes(reorder(Length, counts), counts), fill = "dodgerblue", stat = "identity") +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Length", title = "Number of Fire/Explosion (Non-Fuel) Incidents by Vessel Length", subtitle = "For vessels under 20 feet") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 8))
```
### Fire/Explosion (unknown origin)
Firstly, we'll look into the cause of this type of fire/explosion. However, I don't think we'll gleam much of anything for this entire category...as if the origin is unknown, how would we know the cause? How could we place blame on machinery or electrical problems? Nonetheless, I'll run the code for completeness's sake. 
```{r}
Unknown_related1 <- Boats_lessthan20 %>%
  filter(VesselEvent1=="Fire/explosion (unknown origin)") 

Unknown_related2 <- Boats_lessthan20 %>%
  filter(VesselEvent2=="Fire/explosion (unknown origin)")

Unknown_related3 <- Boats_lessthan20 %>%
  filter(VesselEvent3=="Fire/explosion (unknown origin)")

Unknown_related_under20 <- rbind(Unknown_related1, Unknown_related2, Unknown_related3)


df_VesselCause1_Unknownunder20 <- Unknown_related_under20 %>%
  select(VesselCause1, VesselCause2, VesselCause3)

df_VesselCause2_Unknownunder20 <- Unknown_related_under20 %>%
  select(VesselCause1,VesselCause2, VesselCause3) %>%
  mutate("VesselCause1"=VesselCause2)

df_VesselCause3_Unknownunder20 <- Unknown_related_under20 %>%
  select(VesselCause1,VesselCause3, VesselCause2) %>%
  mutate("VesselCause1"=VesselCause3)

Unknown_related_under20_CauseTally <- rbind(df_VesselCause1_Unknownunder20, df_VesselCause2_Unknownunder20, df_VesselCause3_Unknownunder20)

Unknown_related_under20_CauseTally1 <- Unknown_related_under20_CauseTally %>%
  group_by(VesselCause1) %>%
  filter(VesselCause1!="na") %>%
  summarise(counts = n())

# Now, lets graph it up! Finally..
ggplot(Unknown_related_under20_CauseTally1, aes(x = VesselCause1, y= counts )) +
  geom_bar(aes(reorder(VesselCause1, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Incident Cause", title = "Number of Fire/Explosion (Unknown Origin) Incidents by Causes Listed", subtitle = "For vessels under 20 feet") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 5)) 
```
Now, machinery and equipment failures
Again, I don't expect we'll see any results captured. Unknown origin implies, we don't know..
```{r}
df_VesselMachineryFailure_Unknownunder20 <- Unknown_related_under20 %>%
  group_by(VesselMachineryFailure) %>%
  filter(VesselMachineryFailure!="na") %>%
  summarise(counts = n()) 

ggplot(df_VesselMachineryFailure_Unknownunder20, aes(x = VesselMachineryFailure, y = counts)) +
  geom_bar(aes(reorder(VesselMachineryFailure, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Machinery Failure (Type)", title = "Number of Fire/Explosion (Unknown Origin) Incidents by Machinery Failure Reported", subtitle = "For vessels under 20 feet") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 8))

################################################################################################

df_VesselEquipmentFailure_Unknownunder20 <- Unknown_related_under20 %>%
  group_by(VesselEquipmentFailure) %>%
  filter(VesselEquipmentFailure!="na") %>%
  summarise(counts = n()) 
#As we can see, there were no incidents, where a fire extinguisher failed..
```

What lengths under 20 feet, had fire/explosion (Unknown Origin) related accidents?
This is a category we can gleam more information from. We may not know the origin, but we'll rest assured know the vessel lengths.
```{r}
df_VesselLength_Unknownunder20<- Unknown_related_under20 %>%
  group_by(Length) %>%
  summarise(counts = n()) 

ggplot(df_VesselLength_Unknownunder20, aes(x = Length, y = counts)) +
  geom_bar(aes(reorder(Length, counts), counts), fill = "dodgerblue", stat = "identity") +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Length", title = "Number of Fire/Explosion (Unknown Origin) Incidents by Vessel Length", subtitle = "For vessels under 20 feet") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 5))
```

### Outboard Electrical Problems for Vessels under 20 Feet in Length
```{r}
# This query tells us more information about the __ incidents captured for outboard motor vessels under 20 feet in length. If anyone would like to know more information about these incidents, the data is there to look further into the issue (ex. narrative info)
Electrical_under20 <- Boats_lessthan20 %>%
  filter(VesselMachineryFailure!= "na", VesselMachineryFailure=="Electrical") %>%
  select(BARDID, RegistrationNumber, HIN, VesselMachineryFailure) 

Boats_lessthan20 %>%
  count(OtherMachineryFailure)
#As we can see, none of these list an `other` machinery failure other than what's given as options above. 
```

###Sinking/Swamping/Capsizing
Let's wrangle any cases that have `Capsizing`, `Flooding/swamping`, or `Sinking` listed as the first, second, or third `VesselEvent`. We'll name this group, `Sink_related_under20`
```{r}
Sink_related1 <- Boats_lessthan20 %>%
  filter(VesselEvent1=="Capsizing" | VesselEvent1=="Flooding/swamping" | VesselEvent1=="Sinking") 

Sink_related2 <- Boats_lessthan20 %>%
  filter(VesselEvent2=="Capsizing" | VesselEvent2=="Flooding/swamping" | VesselEvent2=="Sinking")

Sink_related3 <- Boats_lessthan20 %>%
  filter(VesselEvent3=="Capsizing" | VesselEvent3=="Flooding/swamping" | VesselEvent3=="Sinking")

Sink_related_under20 <- rbind(Sink_related1, Sink_related2, Sink_related3)


df_VesselCause1_Sinkunder20 <- Sink_related_under20 %>%
  select(VesselCause1, VesselCause2, VesselCause3)

df_VesselCause2_Sinkunder20 <- Sink_related_under20 %>%
  select(VesselCause1,VesselCause2, VesselCause3) %>%
  mutate("VesselCause1"=VesselCause2)

df_VesselCause3_Sinkunder20 <- Sink_related_under20 %>%
  select(VesselCause1,VesselCause3, VesselCause2) %>%
  mutate("VesselCause1"=VesselCause3)

Sink_related_under20_CauseTally <- rbind(df_VesselCause1_Sinkunder20, df_VesselCause2_Sinkunder20, df_VesselCause3_Sinkunder20)

Sink_related_under20_CauseTally1 <- Sink_related_under20_CauseTally %>%
  group_by(VesselCause1) %>%
  filter(VesselCause1!="na") %>%
  summarise(counts = n())

# Now, lets graph it up! Finally..
# Remember...we have 566 incidents whereby sinking/flooding/capsizing is mentioned for VesselEvent1, 2, or 3. 
# When we look further into the causal breakdown, its going to be greater than 566 in total, because many "VesselEvents" have more than one cause listed. 
ggplot(Sink_related_under20_CauseTally1, aes(x = VesselCause1, y= counts )) +
  geom_bar(aes(reorder(VesselCause1, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Incident Cause", title = "Number of Sink/Swamp/Capsizing Related Incidents by Causes Listed", subtitle = "For vessels under 20 feet") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 200)) 
```

Now, machinery and equipment failures
```{r}
df_VesselMachineryFailure_Sinkunder20 <- Sink_related_under20 %>%
  group_by(VesselMachineryFailure) %>%
  filter(VesselMachineryFailure!="na", VesselMachineryFailure!="Unknown") %>%
  summarise(counts = n()) 

ggplot(df_VesselMachineryFailure_Sinkunder20, aes(x = VesselMachineryFailure, y = counts)) +
  geom_bar(aes(reorder(VesselMachineryFailure, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Machinery Failure (Type)", title = "Number of Sink/Swamp/Capsizing Related Incidents by Machinery Failure Reported", subtitle = "For vessels under 20 feet") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 75))

################################################################################################

df_VesselEquipmentFailure_Sinkunder20 <- Sink_related_under20 %>%
  group_by(VesselEquipmentFailure) %>%
  filter(VesselEquipmentFailure!="na") %>%
  summarise(counts = n()) 
```

What lengths under 20 feet, had sink/flood/capsize related accidents?
```{r}
df_VesselLength_Sinkunder20<- Sink_related_under20 %>%
  group_by(Length) %>%
  summarise(counts = n()) 

ggplot(df_VesselLength_Sinkunder20, aes(x = Length, y = counts)) +
  geom_bar(aes(reorder(Length, counts), counts), fill = "dodgerblue", stat = "identity") +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Length", title = "Number of Sink/Swamp/Capsizing Related Incidents by Vessel Length", subtitle = "For vessels under 20 feet") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 250))
```


## Boats 20-26 feet in length. 
### Fire/Explosion (fuel) Related
First, let's wrangle any cases that have `Fire/explosion (fuel)` listed as the first, second, or third `VesselEvent`. We'll name this group, `Fuel_related_20to26`
```{r}
Fuel_related1a <- Boats_20to26 %>%
  filter(VesselEvent1=="Fire/explosion (fuel)") 

Fuel_related2a <- Boats_20to26 %>%
  filter(VesselEvent2=="Fire/explosion (fuel)")

Fuel_related3a <- Boats_20to26 %>%
  filter(VesselEvent3=="Fire/explosion (fuel)")

Fuel_related_20to26 <- rbind(Fuel_related1a, Fuel_related2a, Fuel_related3a)

# Let's run a few counts to ensure the numbers are adding up
Boats_20to26 %>%
  count(VesselEvent1)

Boats_20to26 %>%
  count(VesselEvent2)

Boats_20to26 %>%
  count(VesselEvent3)
# Ensure the total count for these three above queries is equal toour dataframe, `Fuel_related_20to26`
```

Taking a look at the "Environment" window tells us how many total fire/explosion cases there were related to fuel during this period. 

Now, lets generate some basic descriptive statistics
```{r}
Fuel_related_20to26 %>%
  group_by(VesselCause1, VesselCause2, VesselCause3) %>%
  tally
# While this function isn't perfect, it does allow us to easily count up how often specific causes come up across several pages of incidents.
```

```{r}
# This is another way to accomplish the above code chunk. Count how many instances each VesselCause comes up
df_VesselCause1_fuelunder20<- Fuel_related_under20 %>%
  group_by(VesselCause1) %>%
  summarise(counts = n()) 

df_VesselCause2_fuelunder20<- Fuel_related_under20 %>%
  group_by(VesselCause2) %>%
  summarise(counts = n()) 

df_VesselCause3_fuelunder20<- Fuel_related_under20 %>%
  group_by(VesselCause3) %>%
  summarise(counts = n()) 

# If you summarize the values for each of the three dataframes above, you can get to how many total causes are mentioned for either cause1, cause2, or cause3

```

```{r}
df_VesselCause1_fuel20to26 <- Fuel_related_20to26 %>%
  select(VesselCause1, VesselCause2, VesselCause3)

df_VesselCause2_fuel20to26 <- Fuel_related_20to26 %>%
  select(VesselCause1,VesselCause2, VesselCause3) %>%
  mutate("VesselCause1"=VesselCause2)

df_VesselCause3_fuel20to26 <- Fuel_related_20to26 %>%
  select(VesselCause1,VesselCause3, VesselCause2) %>%
  mutate("VesselCause1"=VesselCause3)

Fuel_related_20to26_CauseTally <- rbind(df_VesselCause1_fuel20to26, df_VesselCause2_fuel20to26, df_VesselCause3_fuel20to26)

Fuel_related_20to26_CauseTally1 <- Fuel_related_20to26_CauseTally %>%
  group_by(VesselCause1) %>%
  filter(VesselCause1!="na") %>%
  summarise(counts = n())

# Now, lets graph it up! Finally..
ggplot(Fuel_related_20to26_CauseTally1, aes(x = VesselCause1, y= counts )) +
  geom_bar(aes(reorder(VesselCause1, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Incident Cause", title = "Number of Fire/Explosion (Fuel) Incidents by Causes Listed", subtitle = "For vessels 20'-26' in length") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 15)) 
```

Now, machinery and equipment failures
```{r}
df_VesselMachineryFailure_fuel20to26 <- Fuel_related_20to26 %>%
  group_by(VesselMachineryFailure) %>%
  filter(VesselMachineryFailure!="na") %>%
  summarise(counts = n()) 

ggplot(df_VesselMachineryFailure_fuel20to26, aes(x = VesselMachineryFailure, y = counts)) +
  geom_bar(aes(reorder(VesselMachineryFailure, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Machinery Failure (Type)", title = "Number of Fire/Explosion (Fuel) Incidents by Machinery Failure Reported", subtitle = "For vessels 20'-26' in length") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 10))

################################################################################################

df_VesselEquipmentFailure_fuel20to26 <- Fuel_related_20to26 %>%
  group_by(VesselEquipmentFailure) %>%
  filter(VesselEquipmentFailure!="na") %>%
  summarise(counts = n()) 
#As we can see, there were no incidents.
```

What lengths 20'-26', had fire/explosion (fuel) related accidents?
```{r}
df_VesselLength_fuel20to26 <- Fuel_related_20to26 %>%
  group_by(Length) %>%
  summarise(counts = n()) 

ggplot(df_VesselLength_fuel20to26, aes(x = Length, y = counts)) +
  geom_bar(aes(reorder(Length, counts), counts), fill = "dodgerblue", stat = "identity") +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Length", title = "Number of Fire/Explosion (Fuel) Incidents by Vessel Length", subtitle = "For vessels 20'-26' in length") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 15))
```

###Fire/Explosion (non-fuel) Related for outboard motor vessels 20'-26' in length 
```{r}
NonFuel_related1a <- Boats_20to26 %>%
  filter(VesselEvent1=="Fire/explosion (non-fuel)") 

NonFuel_related2a <- Boats_20to26 %>%
  filter(VesselEvent2=="Fire/explosion (non-fuel)")

NonFuel_related3a <- Boats_20to26 %>%
  filter(VesselEvent3=="Fire/explosion (non-fuel)")

NonFuel_related_20to26 <- rbind(NonFuel_related1a, NonFuel_related2a, NonFuel_related3a)


df_VesselCause1_Nonfuel20to26 <- NonFuel_related_20to26 %>%
  select(VesselCause1, VesselCause2, VesselCause3)

df_VesselCause2_Nonfuel20to26 <- NonFuel_related_20to26 %>%
  select(VesselCause1,VesselCause2, VesselCause3) %>%
  mutate("VesselCause1"=VesselCause2)

df_VesselCause3_Nonfuel20to26 <- NonFuel_related_20to26 %>%
  select(VesselCause1,VesselCause3, VesselCause2) %>%
  mutate("VesselCause1"=VesselCause3)

NonFuel_related_20to26_CauseTally <- rbind(df_VesselCause1_Nonfuel20to26, df_VesselCause2_Nonfuel20to26, df_VesselCause3_Nonfuel20to26)

NonFuel_related_20to26_CauseTally1 <- NonFuel_related_20to26_CauseTally %>%
  group_by(VesselCause1) %>%
  filter(VesselCause1!="na") %>%
  summarise(counts = n())

# Now, lets graph it up! Finally..
ggplot(NonFuel_related_20to26_CauseTally1, aes(x = VesselCause1, y= counts )) +
  geom_bar(aes(reorder(VesselCause1, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Incident Cause", title = "Number of Fire/Explosion (Non-Fuel) Incidents by Causes Listed", subtitle = "For vessels 20'-26' in lnegth") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 10)) 

# Just to check that its correct...
NonFuel_related_20to26 %>%
  count(VesselCause1)
```

Now, machinery and equipment failures
```{r}
df_VesselMachineryFailure_Nonfuel20to26 <- NonFuel_related_20to26 %>%
  group_by(VesselMachineryFailure) %>%
  filter(VesselMachineryFailure!="na") %>%
  summarise(counts = n()) 

ggplot(df_VesselMachineryFailure_Nonfuel20to26, aes(x = VesselMachineryFailure, y = counts)) +
  geom_bar(aes(reorder(VesselMachineryFailure, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Machinery Failure (Type)", title = "Number of Fire/Explosion (Non-Fuel) Incidents by Machinery Failure Reported", subtitle = "For vessels 20'-26' in length") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 8))

################################################################################################

df_VesselEquipmentFailure_Nonfuel20to26 <- NonFuel_related_20to26 %>%
  group_by(VesselEquipmentFailure) %>%
  filter(VesselEquipmentFailure!="na") %>%
  summarise(counts = n()) 
#As we can see, there were no incidents, where a fire extinguisher failed..
```

What lengths 20'-26' had fire/explosion (non-fuel) related accidents?
```{r}
df_VesselLength_Nonfuel20to26<- NonFuel_related_20to26 %>%
  group_by(Length) %>%
  summarise(counts = n()) 

ggplot(df_VesselLength_Nonfuel20to26, aes(x = Length, y = counts)) +
  geom_bar(aes(reorder(Length, counts), counts), fill = "dodgerblue", stat = "identity") +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Length", title = "Number of Fire/Explosion (Non-Fuel) Incidents by Vessel Length", subtitle = "For vessels 20'-26' in length") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 8))
```

### Fire/Explosion (unknown origin)
Firstly, we'll look into the cause of this type of fire/explosion. However, I don't think we'll gleam much of anything for this entire category...as if the origin is unknown, how would we know the cause? How could we place blame on machinery or electrical problems? Nonetheless, I'll run the code for completeness's sake. 
```{r}
Unknown_related1a <- Boats_20to26 %>%
  filter(VesselEvent1=="Fire/explosion (unknown origin)") 

Unknown_related2a <- Boats_20to26 %>%
  filter(VesselEvent2=="Fire/explosion (unknown origin)")

Unknown_related3a <- Boats_20to26 %>%
  filter(VesselEvent3=="Fire/explosion (unknown origin)")

Unknown_related_20to26 <- rbind(Unknown_related1a, Unknown_related2a, Unknown_related3a)


df_VesselCause1_Unknown20to26 <- Unknown_related_20to26 %>%
  select(VesselCause1, VesselCause2, VesselCause3)

df_VesselCause2_Unknown20to26 <- Unknown_related_20to26%>%
  select(VesselCause1,VesselCause2, VesselCause3) %>%
  mutate("VesselCause1"=VesselCause2)

df_VesselCause3_Unknown20to26 <- Unknown_related_20to26 %>%
  select(VesselCause1,VesselCause3, VesselCause2) %>%
  mutate("VesselCause1"=VesselCause3)

Unknown_related_20to26_CauseTally <- rbind(df_VesselCause1_Unknown20to26, df_VesselCause2_Unknown20to26, df_VesselCause3_Unknown20to26)

Unknown_related_20to26_CauseTally1 <- Unknown_related_20to26_CauseTally %>%
  group_by(VesselCause1) %>%
  filter(VesselCause1!="na") %>%
  summarise(counts = n())

# Now, lets graph it up! Finally..
ggplot(Unknown_related_20to26_CauseTally1, aes(x = VesselCause1, y= counts )) +
  geom_bar(aes(reorder(VesselCause1, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Incident Cause", title = "Number of Fire/Explosion (Unknown Origin) Incidents by Causes Listed", subtitle = "For vessels 20'-26' in lnegth") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 8)) 
```

Now, machinery and equipment failures
Again, I don't expect we'll see any results captured. Unknown origin implies, we don't know..
```{r}
df_VesselMachineryFailure_Unknown20to26 <- Unknown_related_20to26 %>%
  group_by(VesselMachineryFailure) %>%
  filter(VesselMachineryFailure!="na") %>%
  summarise(counts = n()) 

ggplot(df_VesselMachineryFailure_Unknown20to26, aes(x = VesselMachineryFailure, y = counts)) +
  geom_bar(aes(reorder(VesselMachineryFailure, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Machinery Failure (Type)", title = "Number of Fire/Explosion (Unknown Origin) Incidents by Machinery Failure Reported", subtitle = "For vessels 20'-26' in length") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 6))

################################################################################################

df_VesselEquipmentFailure_Unknown20to26 <- Unknown_related_20to26 %>%
  group_by(VesselEquipmentFailure) %>%
  filter(VesselEquipmentFailure!="na") %>%
  summarise(counts = n()) 
#As we can see, there were no incidents, where a fire extinguisher failed..
```

What lengths 20'-26', had fire/explosion (Unknown Origin) related accidents?
This is a category we can gleam more information from. We may not know the origin, but we'll rest assured know the vessel lengths.
```{r}
df_VesselLength_Unknown20to26 <- Unknown_related_20to26 %>%
  group_by(Length) %>%
  summarise(counts = n()) 

ggplot(df_VesselLength_Unknown20to26, aes(x = Length, y = counts)) +
  geom_bar(aes(reorder(Length, counts), counts), fill = "dodgerblue", stat = "identity") +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Length", title = "Number of Fire/Explosion (Unknown Origin) Incidents by Vessel Length", subtitle = "For vessels 20'-26' in length") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 5))
```

### Outboard Electrical Problems for Vessels 20'-26'  in Length
```{r}
# This query tells us more information about the __ incidents captured for outboard motor vessels under 20 feet in length. If anyone would like to know more information about these incidents, the data is there to look further into the issue (ex. narrative info)
Electrical_20to26 <- Boats_20to26 %>%
  filter(VesselMachineryFailure!= "na", VesselMachineryFailure=="Electrical") %>%
  select(BARDID, RegistrationNumber, HIN, VesselMachineryFailure) 

Boats_20to26 %>%
  count(OtherMachineryFailure)
#As we can see, none of these list an `other` machinery failure other than what's given as options above. 
```

###Sinking/Swamping/Capsizing
Let's wrangle any cases that have `Capsizing`, `Flooding/swamping`, or `Sinking` listed as the first, second, or third `VesselEvent`. We'll name this group, `Sink_related_20to26`
```{r}
Sink_related1a <- Boats_20to26 %>%
  filter(VesselEvent1=="Capsizing" | VesselEvent1=="Flooding/swamping" | VesselEvent1=="Sinking") 

Sink_related2a <- Boats_20to26 %>%
  filter(VesselEvent2=="Capsizing" | VesselEvent2=="Flooding/swamping" | VesselEvent2=="Sinking")

Sink_related3a <- Boats_20to26 %>%
  filter(VesselEvent3=="Capsizing" | VesselEvent3=="Flooding/swamping" | VesselEvent3=="Sinking")

Sink_related_20to26 <- rbind(Sink_related1a, Sink_related2a, Sink_related3a)


df_VesselCause1_Sink20to26 <- Sink_related_20to26 %>%
  select(VesselCause1, VesselCause2, VesselCause3)

df_VesselCause2_Sink20to26 <- Sink_related_20to26 %>%
  select(VesselCause1,VesselCause2, VesselCause3) %>%
  mutate("VesselCause1"=VesselCause2)

df_VesselCause3_Sink20to26 <- Sink_related_20to26 %>%
  select(VesselCause1,VesselCause3, VesselCause2) %>%
  mutate("VesselCause1"=VesselCause3)

Sink_related_20to26_CauseTally <- rbind(df_VesselCause1_Sink20to26, df_VesselCause2_Sink20to26, df_VesselCause3_Sink20to26)

Sink_related_20to26_CauseTally1 <- Sink_related_20to26_CauseTally %>%
  group_by(VesselCause1) %>%
  filter(VesselCause1!="na") %>%
  summarise(counts = n())

# Now, lets graph it up! Finally..
# Remember...we have 566 incidents whereby sinking/flooding/capsizing is mentioned for VesselEvent1, 2, or 3. 
# When we look further into the causal breakdown, its going to be greater than 566 in total, because many "VesselEvents" have more than one cause listed. 
ggplot(Sink_related_20to26_CauseTally1, aes(x = VesselCause1, y= counts )) +
  geom_bar(aes(reorder(VesselCause1, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Incident Cause", title = "Number of Sink/Swamp/Capsizing Related Incidents by Causes Listed", subtitle = "For vessels 20'-26' in length") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 200)) 
```

Now, machinery and equipment failures
```{r}
df_VesselMachineryFailure_Sink20to26 <- Sink_related_20to26 %>%
  group_by(VesselMachineryFailure) %>%
  filter(VesselMachineryFailure!="na") %>%
  summarise(counts = n()) 

ggplot(df_VesselMachineryFailure_Sink20to26, aes(x = VesselMachineryFailure, y = counts)) +
  geom_bar(aes(reorder(VesselMachineryFailure, counts), counts), fill = "dodgerblue", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Machinery Failure (Type)", title = "Number of Sink/Swamp/Capsizing Related Incidents by Machinery Failure Reported", subtitle = "For vessels 20'-26' in length") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 75))

################################################################################################

df_VesselEquipmentFailure_Sink20to26 <- Sink_related_20to26 %>%
  group_by(VesselEquipmentFailure) %>%
  filter(VesselEquipmentFailure!="na") %>%
  summarise(counts = n()) 
```

What lengths 20'-26' feet, had sink/flood/capsize related accidents?
```{r}
df_VesselLength_Sink20to26<- Sink_related_20to26 %>%
  group_by(Length) %>%
  summarise(counts = n()) 

ggplot(df_VesselLength_Sink20to26, aes(x = Length, y = counts)) +
  geom_bar(aes(reorder(Length, counts), counts), fill = "dodgerblue", stat = "identity") +
    labs(y = "Number of Vessel Incidents", fill="test", x = "Length", title = "Number of Sink/Swamp/Capsizing Related Incidents by Vessel Length", subtitle = "For vessels 20'-26' in length") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  scale_y_continuous(limits = c(0, 250))
```

##Moving onto Pontoon Boat Analysis...
```{r}

```

